### 2024-04-30 thesis/chapters/02-analysis.tex
Analysis finished - page 37

\section{Condition monitoring} \label{section:condition-monitoring}
All rotating machinery eventually fails because of the long-term strain on the individual parts, inadequate workmanship, installation, or operational procedures. In the end, these factors cause the equipment not to fulfill its intended functionality. Many instrumentation methods are practiced to reveal evolving faults: vibration and acoustic noise monitoring, electric supply line measurements, thermography, oil and particle analysis, ultrasonic testing, etc. Vibration signals are the preferred tool for rotating machinery monitoring \cite{mohanty_machinery_2015}.

The defect needs to be repaired or replaced, preferably without significant production downtime, further damage to the other attached elements, or any endangerment of the responsible personnel. The maintenance strategies are chosen according to the machine's importance as a result of potential failure's impact on the system. The guide to set appropriate maintenance procedures as outlined in the IEC~60706-2 standard, and involves reliability-centered maintenance analysis~\cite{el-thalji_predictive_2019}.

\subsection{Maintenance strategies}
There are three different approaches to maintenance across the industry: \textbf{reactive, preventive, and predictive}~\cite{scheffer_practical_2004}. In general, the more sophisticated methods are beneficial in a high-stakes environment. The unexpected machine shutdown can have a negative economic impact on the enterprise, resulting in decreased product quality and demands for spare parts to be ready in the supply inventory at all times. In certain situations, it suffices to utilize a simpler maintenance program, but predictive maintenance gains attraction in Industry 4.0 to optimize usage of assets~\cite{cinar_machine_2020}.

\textbf{Reactive maintenance} allows machinery to run until a complete failure. This is the most inappropriate way to maintain the production line, but it is straightforward enough. It requires a large stock of replacement parts on-site and breakage inflicts a ``crisis management mode'' upon the plant \cite{scheffer_practical_2004}. On-demand repairs are justified when short downtime is acceptable, full and swift replacement of a broken machine with a backup is possible, or there is a negligible threat to the surrounding environment from failure~\cite{ziaran_technicka_2013}.

\textbf{Preventive maintenance} is performed before any issue is detected. Maintenance occurs at regular intervals derived from a predetermined period in the calendar or expected machine running time (MTTF - Mean Time To Failure). The schedule is crucial but can result in components being replaced in good condition, creating waste. The parts can occasionally stay in operation too long, and the machine breaks as a result. Conservative planning is usually the norm to keep the machines always in a perfect state, and therefore more frequent interventions are required~\cite{mohanty_machinery_2015}.

\textbf{Predictive maintenance} known as condition-based maintenance (CbM), improves the predictability of reactive maintenance and eliminates waste in overall resource utilization of cautious prevention. The machine downtime is scheduled after the detection of unhealthy trends in fault monitoring with sensors and the identification of troublesome components.

@@

The \emph{P-F curve} is a widespread representation of equipment degradation over time based on historical records (Fig.~\ref{fig:p-f-curve}). Corrective action should be taken between the event of potential failure (P), when the fault detection is activated, and functional failure point (F) in the P-F interval~\cite{bousdekis_enterprise_2021}.  These division points are not exactly set but have a statistical distribution to them.

The \emph{Remaining Useful Life} (RUL) of the specific running machine in the given instance can be merely estimated analytically, with the survival probabilities of the individual components, based on the model of the ``run-to-failure'' histories, and usage parameters~\cite{okoh_overview_2014}. Predictive condition monitoring aims to extend lifespan to the maximum by predicting expected RUL.

@@

A high failure rate is present not only at the worn-out stage when the parts are fatigued or corroded, but also in the early stages soon after assembly. Manufacturing or material defects, inadequate installation, or improper start-up procedures, are all suspected causes. During the stable middle phase, malfunction can occur after the machine's excessive overload. The time plot to failure rate is known as the bathtub curve~(Fig.~\ref{fig:bathtub-curve}).

@@

The cause of vibration comes out of the changing force in its magnitude or direction. The most emerging defects can be encompassed by explaining the deficiencies of the mechanical structure. These defects are broadly categorized as \textbf{unbalance, misalignment, looseness, eccentricity, deformation, crack, and influence of the external force} (friction)~\cite{davies_handbook_2012}. It is important to stress that our concern is not the underlying deformities in mechanical parts, but the correct fault classification based on the signal waveform.

@@

One of the methods vibration experts utilize in the identification of the damaged part, according to the frequency spectrum, is \textbf{order analysis}. The excessive peaks at harmonic frequencies are of interest. Harmonics are integer multiples of fundamental frequency (1x rpm)~(Tab.~\ref{tab:vibration-causes}):

@@

%TODO bearing image / equations

@@

Criteria introduced to judge vibration severity are absolute vibration magnitude, change in the magnitude vector, and rate of change. In terms of maximal magnitudes, the machines of varying sizes are split into four severity zones defined in the chart ~(Tab.~\ref{tab:iso20816-vibration-severity}). The values in this table serve as guidelines towards realistic requirements between machine manufacturers and their customers.

@@

The operational limits in the form of \emph{alarms} and \emph{trips} are usually established on the zone boundaries or close to them. Alarms are placed between zones B and C providing a warning about reaching the threshold significant for noticeable change. Trips in between zones C and D urge immediate action or machine shut down. Both limits should not exceed 1.25 times the upper boundary of the lower zones and initially are set based on previous experience with the machine~\cite{noauthor_iso_2002}.

\textbf{ISO 13373-1:2002} delves into further nuances of vibration monitoring and expands on procedures outlined in terms of the vocabulary in ISO 20186. According to the standard, the data collection operates in continuous or periodic observation modes which follow an event or interval. Both designs can be permanently mounted. In continuous collection, it is recommended to have a ``multiplexing rate sufficiently rapid, so there is no significant data or trends lost''~\cite{noauthor_iso_2002}. When channels are scanned in succession with gaps between data points, the system is known as ``scanning''.

@@

The transducer of choice for condition monitoring is the accelerometer, which can provide the acceleration value of the body and velocity after signal integration. However, standard advise against double integrating for displacement. The recommended frequency range for an accelerometer is 0.1~Hz to 30~kHz. The choice of transducer mount significantly lowers its resonance frequency, which is least influenced by the stud mount and stiff cement mount. The resonance is reduced to around 8~kHz with the use of soft epoxy or permanent magnet.

@@

\section{Signal preprocessing} \label{section:signal-preprocessing}
The vibration signals in a factory environment are inherently full of disturbances. Nearby equipment operation and handling of heavy objects in the surroundings can all contribute to the unwanted chaotic movement in otherwise mostly pure oscillatory motion. In addition, accelerometers suffer from systematic measurement errors in the form of thermal noise, zero-g offset as a result of slight miscalibration, and bias originating from a constant force of gravity. These unavoidable distortions are to some extent suppressable by digital filters. In the preprocessing stage, we consider trend removal, noise reduction with adaptive filters, and time synchronous averaging to eliminate external interference.

@@

A steeper 3~dB attenuation band can be achieved by increasing the order of the filter. Then the cutoff frequency should be such that filter coefficients are fractions to counteract rounding errors~\cite{tittelbach-helmrich_digital_2021}.

@@

Predictive maintenance has ideal prerequisites for the application of feature engineering because the signal is usually pseudo-stationary, and the trend monitoring variables come out of extensive domain expertise in mechanics. The advantages of the add-in extraction effort, as opposed to processing samples unmodified, are to gain better classification precision, reduce computational burden and storage capacity downstream with dimensionality reduction~\cite{johnson_feature_2019}.

It is important to note that the design of features is not a standalone step in the machine learning pipeline, but it should be done iteratively to improve the target model. Signal features are computed in the time and frequency domain~\cite{brito_fault_2021}.

@@

The other significant time domain attributes are derived as ratios of previous simpler ones. These ratios are crest, clearance, impulse factor, and shape factor~(Tab.~\ref{tab:td-features})~\cite{nandi_condition_2019}. Many articles have been successful in fault detection of bearings out of transients in impulsive signals with kurtosis, crest, and clearance factor~\cite{brito_fault_2021}. It is also suggested that the shape factor can signify unbalance and misalignment faults~\cite{nandi_condition_2019}.

@@

In the frequency domain, we can obtain the usual statistical properties of the distribution being spectral centroid, skew and kurtosis. Additionally, spectral roll-on and roll-off, fundamental frequency, entropy, negentropy, spectral flux, signal-to-noise ratio (noisiness), energy in frequency bands, and energy ratio are extracted (Tab.~\ref{tab:fd-features})~\cite{peeters_large_2004}.

In geometric terms, the spectral centroid represents the barycenter of the frequency magnitude plot. Spectral roll-off gives a notion about the spectral distribution because it identifies the frequency $f_c$ below which 95\% of the signal energy is contained. Complementary, the roll-on frequency is chosen so that 5\% of the signal energy is below this value. According to the definition, spectral flux is normalized cross-correlation between two successive amplitude spectra a value of one means the spectra are the most dissimilar.

@@

The essential peak-finding approaches are based either on magnitude or gradient. All found extrema are commonly filtered with the magnitude of prominences and the width at half prominence. In the magnitude-based method, the middle point $x_i$ is compared to two neighboring points and the peak is then: $x_{i-1} < x_i > x_{i+1}$. The gradient-based method evaluates the first derivative at the point which is equal to zero if the point is a local maximum, local minimum, or inflection point~\cite{adikaram_non-parametric_2016}.

@@

\subsection{Bearing defect frequencies}

%TODO n = number of balls, d - average of outer ring diameter and outer ring, f_s - rotation speed, beta - contact angle - ball bearings close to 0

% Rolling element bearings have four very important components, two circular races, between which the rolling elements—rollers, tapered rollers, or spherical balls—are held. The fourth important element in a rolling element bearing is the cage or the separator. The cage ensures that no two rolling elements come in contact with each other
% The bearing vibration is further complicated when the radial loads on the bearing change during rotation of the shaft, and defects like impurities or scratches, pits, and so on, develop on the bearing element surfaces
% The characteristic bearing defect frequencies for the different rolling elements of bearings can be determined

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Feature}            & \textbf{Equation}                                                                   
 \\ \hline
\textit{Outer race fault frequency} & $ BPFO = f_s \cdot \frac{n}{2} \cdot \left(1 - \frac{d}{D}\cos \beta \right)$               
\\ \hline
\textit{Inner race fault frequency} & $ BPFI = f_s \cdot \frac{n}{2} \cdot \left(1 + \frac{d}{D}\cos \beta \right)$
\\ \hline
\textit{Cage rotation frequency}    & $ FTF = f_s \cdot \frac{1}{2} \cdot \left(1 - \frac{d}{D}\cos \beta \right) $      
\\ \hline
\textit{Ball defect frequencies} & $ BSF = f_s \cdot \frac{D}{d} \cdot \left(1 + \left(\frac{D}{d}\cos \beta\right)^2 \right) $
\\ \hline
\end{tabular}
\caption{Bearing defect frequencies}
\label{tab:bearing-defect-features}
\end{table}

\subsection{Time-frequency domain features}
The \textbf{Short-time Fourier transform} (STFT) splits the time-domain signal into equal-length intervals. Individual chunks have a 50\% overlap and are multiplied with weights of window function to balance scalloping loss and spectral leakage due to Fourier transform periodicity assumption. Traditionally, the Hann window is commonly used instead of the rectangular window~\cite{ziaran_technicka_2013,noauthor_iso_2016_2}.

@@

In washing machine status classification, the discrete wavelet transform with Daubechies wavelet (db4) and fifth-level decomposition provide features with a combination of approximation (cA5) and detail coefficients (cD1, \dots, cD5). Washing machines belong to three categories: no fault, electric motor clamping screws problem, and a loose or broken counterweight. Extracted measures were sample mean and sample variances over autocorrelation functions of coefficients (AcD$n$) and smoothed coefficients cD1, cD2 by moving average filter~\cite{goumas_classification_2002}.

@@

Tool wear diagnosis based on acoustic emission (AE) signal considers \emph{wavelet packet energy} in bands $E_{8}$, $E_{10}$, $E_{12}$, \emph{energy ratios} $P_{8}$, $P_{13}$, and \emph{energy entropy} as having high correlation ($|r| > 0.8$) with the band saw flank face width. The acoustic signal is decomposed into three layers using Daubechies db3 wavelet. The bottom layer contains bands numbered 7 through 14, each with a bandwidth of 62.5 kHz, because of the 1 MHz sampling frequency. The feature vector constructed in the article includes other statistical metrics out of power spectral density that has reached a notable correlation with evolving wear. The statistics are \emph{skewness}, \emph{kurtosis}, \emph{shape factor}, and \emph{centroid frequency}~\cite{zhuo_research_2022}.

@@

Wavelet domain features and associated spectral segmentation appear to be a powerful tool in conserving data bandwidth. These variables turn out to be difficult to communicate and comprehend. Therefore, we will use more standardized metrics in further studies.

@@

\item \textbf{Correlation coefficient} - expresses the linear relationship between two variables. The codependent variables are of three sorts: quantitative, ordinal, and nominal. The choice of coefficient calculation is determined by the type of variables under consideration shown in Table \ref{tab:corr-coef}.

@@

\item \textbf{Mutal information} - quantifies the dependence between features, or between features and class labels. It is almost identical to Information Gain. The probability distribution of proximity of variables derives from the relative entropy known as the Kullback-Leibler distance. Mutual information presumes variables are discrete. In the case of quantitative variables, mutual information is estimated by binning or nearest neighbors methods~\cite{ross_mutual_2014}.

@@

The measurements coming in the steaming fashion have to be processed in a single pass. The detection model must deal with the minimal admissible assumptions about the nature of the input events. The outliers are derived based on non-parametric statistical models, nearest-neighbor clustering, and isolation-based approaches~\cite{gervasi_anomaly_2020}.
\bigbreak

\textbf{DenStream} is a density-based algorithm adapted from DBSCAN to cluster streaming data of arbitrarily shaped groups. Samples it includes in the first step into coherent clusters are core data points in each other's neighborhoods. Core points have at least \emph{MinPts} ($\mu$) points in their neighborhood of radius \emph{Eps} ($\varepsilon$) units. Then non-core points in the proximity area of the core point are attached to the cluster containing it~\cite{aggarwal_data_2014}.

Quality of clustering results is evaluated by \emph{Silhouette score} in the range [-1; 1]. It demands points within clusters to have high cohesion and at the same time to have large separation from other clusters. The low score indicates a too small or too large number of clusters~\cite{rousseeuw_rousseeuw_1987}.

@@

In the online maintenance phase, DenStream summarizes the nearby observations into core \emph{micro-clusters} that can be potential micro-clusters or outlier \emph{micro-clusters} (Fig.~\ref{fig:denstream})~\cite{ghesmoune_state---art_2016}. The (outlier) \emph{o-micro-clusters} can grow into (potential) \emph{p-micro-clusters} when they encompass $\beta \mu$ points. The outliers are discounted after some time in accordance to the decay function: $f(t) = 2^{-\lambda t}$ or below the lower weight limit $\xi$. The on-demand offline stage runs DBSCAN over the approximate representation in micro-clusters to deliver final apportionment~\cite{cao_density-based_2006}.
\bigbreak

%\cite{tan_fast_2011}.

\subsection{Classification}
Accurate multi-class classification of machine fault causes according to the characteristics of known ones is a much more difficult task than novelty detection. Fault combinations have to be recorded and transformed into feature space. Interactions among fault root causes have to be considered. We are aware of rapid advances in knowledge transfer for deep neural networks~\cite{maurya_condition-based_2021}. So far, solutions seem not production-ready and hard to explain. Therefore, we opt to use a simpler model.

Performance of classification is estimated by several metrics on the validation set obtained using hold-out or cross-validation techniques. Frequently used quantities for classifier model evaluation include accuracy, precision, recall, f1 score, area under the ROC curve, and counts of hits and misses in a confusion matrix.
\bigbreak

\textbf{K-nearest neighbors} (k-NN) assigns the data point to the class where the majority of $k$ closest instances belong (Fig.~\ref{fig:KNN}). It means it can work in a semi-supervised environment because it can infer labels just from knowing a few annotations. The major drawback of k-NN is a preference for the majority class in imbalanced class-size datasets~\cite{shi_improving_2020}. The issue is mitigated with class weights or resampling classes by oversampling or undersampling. The k-NN algorithm requires features to be normalized to assign the same importance to each predictor.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/analysis/KNN.png}
        \caption{k-NN with k = 5}
        \label{fig:KNN}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/analysis/M-tree.png}
        \caption{M-tree data structure}
        \label{fig:m-tree}
    \end{subfigure}
    \caption{Nearest neighbors classification algorithm~\cite{chen_skyline_2009}}
\end{figure}

The sense of distance between feature vectors $\mathbf{x}$, $\mathbf{y}$ have to be defined, so several metrics are available like \emph{Euclidian distance}, \emph{Mahalanobis distance}, \emph{Manhatann distance} etc. (Tab.~\ref{tab:KNN-distance})~\cite{sheng_review_2020, abu_alfeilat_effects_2019}. The optimal $k$ parameter is set by supervised learning according to the breaking point in the elbow curve that plots choices of $k$ against the error rate. The demanding neighborhood queries are sped up using spatial index in spatial databases that utilizes search tree such as kd-tree, R-tree, or M-tree.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|}
\hline
\textbf{Distance}     & \textbf{$d(\mathbf{x}, \mathbf{y})$}                                   \\ \hline
Manhatann distance	 & $ |x_i - y_i| $													   \\ \hline
Euclidian distance    & $ \sqrt{\sum_{i = 1}^{n}(x_i - y_i)^2} $                               \\ \hline
Mahalanobis distance  & $ (\mathbf{x} - \mathbf{y})^T C^{-1} (\mathbf{x} - \mathbf{y}) $       \\ \hline
\end{tabular}
\caption{Distance metrics for k-NN}
\label{tab:KNN-distance}
\end{table}

>>>>

\section{Condition monitoring} \label{section:condition-monitoring}
All rotating machinery eventually fails because of the long-term strain on the individual parts, inadequate workmanship, installation, or operational procedures. In the end, these factors cause the equipment not to fulfill its intended functionality. Many instrumentation methods are practiced to reveal evolving faults: vibration and acoustic noise monitoring, electric supply line measurements, thermography, oil and particle analysis, ultrasonic testing, etc. Vibration signals are the preferred tool for rotating machinery monitoring~\cite{mohanty_machinery_2015}.

The defect needs to be repaired or replaced, preferably without significant production downtime, further damage to the other attached elements, or any endangerment of the responsible personnel. The maintenance strategies are chosen according to the machine's importance as a result of potential failure's impact on the system. The guide to set appropriate maintenance procedures is outlined in the IEC~60706-2 standard and involves reliability-centered maintenance analysis~\cite{el-thalji_predictive_2019}.

\subsection{Maintenance strategies}
There are three different approaches to maintenance across the industry: \textbf{reactive, preventive, and predictive}~\cite{scheffer_practical_2004}. In general, the more sophisticated methods are beneficial in a high-stakes environment. The unexpected machine shutdown can have a negative economic impact on the enterprise which results in decreased product quality and demands for spare parts to be ready in the supply inventory at all times. In certain situations, it suffices to utilize a simpler maintenance program, but predictive maintenance gains attraction in Industry 4.0 to optimize usage of assets~\cite{cinar_machine_2020}.

\textbf{Reactive maintenance} allows machinery to run until a complete failure which is the most inappropriate way to maintain the production line, but it is straightforward enough. It requires a large stock of replacement parts on-site. Any breakage inflicts crisis management upon the plant \cite{scheffer_practical_2004}. On-demand repairs are justified when short downtime is acceptable, full and swift replacement of a broken machine with a backup is possible, or there is a negligible threat to the surrounding environment from failure~\cite{ziaran_technicka_2013}.

\textbf{Preventive maintenance} is performed before any issue is detected. Maintenance occurs at regular intervals derived from a predetermined period in the calendar or expected machine running time and its mean time to failure (MTTF). The schedule is crucial but can result in components being replaced in good condition, creating waste. The parts can occasionally stay in operation too long, and the machine breaks as a result. Conservative planning is usually the norm to keep the machines in perfect shape. Therefore, more frequent interventions are required~\cite{mohanty_machinery_2015}.

\textbf{Predictive maintenance} (PdM) known also as condition-based maintenance (CbM), improves the predictability of reactive maintenance and eliminates waste in overall resource utilization of cautious prevention. The machine downtime is scheduled after the detection of unhealthy trends in fault monitoring with sensors and the identification of troublesome components.

@@

The \emph{P-F curve} is a widespread representation of equipment's degradation over time based on historical records (Fig.~\ref{fig:p-f-curve}). Corrective action should be taken between the event of potential failure (P), when the fault detection is activated, and functional failure point (F) in the P-F interval~\cite{bousdekis_enterprise_2021}. These division points are not set precisely but have a statistical distribution.

The \emph{Remaining Useful Life} (RUL) of the specific running machine in the given instance can be merely estimated analytically, with the survival probabilities of the individual components, based on the model of the ``run-to-failure'' history and usage parameters~\cite{okoh_overview_2014}. Predictive condition monitoring aims to extend the lifespan to the maximum by predicting expected RUL.

@@

A high failure rate is present at the worn-out stage when the parts are fatigued or corroded. It can also be registered in the early stages soon after assembly. Manufacturing or material defects, inadequate installation, or improper start-up procedures are all suspected causes. During the stable middle phase, malfunction can occur after the machine's excessive overload. The time plot to failure rate is known as the bathtub curve~(Fig.~\ref{fig:bathtub-curve}).

@@

The cause of vibration comes out of the changing force in its magnitude or direction. The most emerging defects can be encompassed by explaining the deficiencies of the mechanical structure. These defects are broadly categorized as \textbf{unbalance, misalignment, looseness, eccentricity, deformation, crack, and influence of the external force} (friction)~\cite{davies_handbook_2012}. It is important to stress that our concern are not the underlying deformities in mechanical parts but the correct fault classification based on the signal waveform.

@@

One of the methods vibration experts utilize in the identification of the damaged part, according to the frequency spectrum, is \textbf{order analysis}. The excessive peaks in harmonic frequencies are of interest. Harmonics are integer multiples of fundamental frequency (1x rpm)~(Tab.~\ref{tab:vibration-causes}):

@@

The mechanical properties of bearings are instrumental in identifying the characteristic frequencies of their components~\cite{mohanty_machinery_2015, ziaran_technicka_2013}. Bearings are composed of outer and inner circular races between which the rolling elements are held inside the cage to avoid mutual contact~(Fig.\ref{fig:bearing-parts}). The developing scratches or pits can appear as increases in amplitude around characteristic frequencies and their harmonics but sometimes this is not enough for accurate diagnostics~\cite{brito_fault_2021}.

The table \ref{tab:bearing-defect-features} provides formulas for rolling bearing defect frequencies at rotational speed $f_s$ where $n$ is a number of rolling elements, $d$ is the average of outer and inner ring diameters, and contact angle $\beta$. Because the frequencies of components depend on shaft rotational speed the values are supplied as its multiplication factors alongside datasets.

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Feature}            & \textbf{Equation}                                                                   
 \\ \hline
\textit{Outer race fault frequency} & $ BPFO = f_s \cdot \frac{n}{2} \cdot \left(1 - \frac{d}{D}\cos \beta \right)$               
\\ \hline
\textit{Inner race fault frequency} & $ BPFI = f_s \cdot \frac{n}{2} \cdot \left(1 + \frac{d}{D}\cos \beta \right)$
\\ \hline
\textit{Cage rotation frequency}    & $ FTF = f_s \cdot \frac{1}{2} \cdot \left(1 - \frac{d}{D}\cos \beta \right) $      
\\ \hline
\textit{Ball defect frequencies} & $ BSF = f_s \cdot \frac{D}{d} \cdot \left(1 + \left(\frac{D}{d}\cos \beta\right)^2 \right) $
\\ \hline
\end{tabular}
\caption{Characteristic frequencies of bearings}
\label{tab:bearing-defect-features}
\end{table}

@@

Criteria introduced to judge vibration severity are absolute vibration magnitude, change in the magnitude vector, and rate of change. In terms of maximal magnitudes, the machines of varying sizes are split into four severity zones defined in the chart~(Tab.~\ref{tab:iso20816-vibration-severity}). The values in this table serve as guidelines for realistic requirements between machine manufacturers and their customers.

@@

The operational limits in the form of \emph{alarms} and \emph{trips} are usually established on the zone boundaries or close to them. Alarms are placed between zones B and C providing a warning about reaching the threshold significant for noticeable change. Trips in between zones C and D urge immediate action or machine shut down. Both limits should not exceed 1.25 times the upper boundary of the lower zones and are initially set based on previous experience with the machine~\cite{noauthor_iso_2002}.

\textbf{ISO 13373-1:2002} delves into further nuances of vibration monitoring and expands on procedures outlined in terms of the ISO 20186 vocabulary. According to the standard, the data collection operates in continuous or periodic observation modes that follow an event or interval. Both designs can be permanently mounted. In continuous collection, it is recommended to have a ``multiplexing rate sufficiently rapid, so there is no significant data or trends lost''~\cite{noauthor_iso_2002}. When channels are scanned in succession with gaps between data points, the system is known as ``scanning''.

@@

The transducer of choice for condition monitoring is the accelerometer. It measures the acceleration value of the body and velocity after signal integration. However, standards advise against double integrating for displacement. The recommended frequency range for an accelerometer is 0.1~Hz to 30~kHz. The choice of transducer mount significantly influences its resonance frequency. The stud mount and stiff cement mount least lower the effective resonance. The resonance is reduced to around 8~kHz when using the soft epoxy or permanent magnet.

@@

\section{Signal preprocessing} \label{section:signal-preprocessing}
The vibration signals in a factory environment are inherently full of disturbances. Nearby equipment operation and heavy objects handling in the surroundings can all contribute to the unwanted chaotic movement in otherwise mostly pure oscillatory motion. In addition, accelerometers suffer from systematic measurement errors in the form of thermal noise, zero-g offset from slight miscalibration, and bias originating from a constant force of gravity. These unavoidable distortions are somewhat suppressable by digital filters. In the preprocessing stage, we consider trend removal and time synchronous averaging to eliminate external interference.

@@

A steeper 3~dB attenuation band is achieved by increasing the order of the filter. Then, the cutoff frequency should be such that filter coefficients are fractions to counteract rounding errors~\cite{tittelbach-helmrich_digital_2021}.

@@

Predictive maintenance has ideal prerequisites for the application of feature engineering because the signal is usually pseudo-stationary, and the trend monitoring variables come out of extensive domain expertise in mechanics. The advantages of the add-in extraction effort as opposed to processing unmodified samples are to gain better classification precision and reduce computational burden and storage capacity downstream with dimensionality reduction~\cite{johnson_feature_2019}.

It is important to note that the design of features is not a standalone step in the machine learning pipeline, but it should be performed iteratively to improve the target model. Signal features are computed in the time and frequency domain~\cite{brito_fault_2021}.

@@

The other significant time-domain attributes are derived as ratios of previous simpler ones. These ratios are crest, clearance, impulse factor, and shape factor~(Tab.~\ref{tab:td-features})~\cite{nandi_condition_2019}. Many articles have been successful in fault detection of bearings out of transients in impulsive signals with kurtosis, crest, and clearance factor~\cite{brito_fault_2021}. It is also suggested that the shape factor can signify unbalance and misalignment faults~\cite{nandi_condition_2019}.

@@

In the frequency domain, we can obtain the usual statistical properties of the distribution which are spectral centroid, skew, and kurtosis. Additionally, spectral roll-on and roll-off, fundamental frequency, entropy, negentropy, spectral flux, signal-to-noise ratio (noisiness), energy in frequency bands, and energy ratio are extracted (Tab.~\ref{tab:fd-features})~\cite{peeters_large_2004}.

In geometric terms, the spectral centroid represents the barycenter of the frequency magnitude plot. Spectral roll-off gives a notion about the spectral distribution because it identifies the frequency $f_c$ below which 85\% of the signal energy is contained. The roll-on frequency is chosen so that 5\% of the signal energy is below this value. According to the definition, spectral flux is normalized cross-correlation between two successive amplitude spectra. A flux value of one means the spectra are the most dissimilar.

@@

The essential peak-finding approaches are based either on magnitude or gradient. All found extrema are commonly filtered with the magnitude of prominences and the width at half prominence. In the magnitude-based method, the middle point $x_i$ is compared to two neighbouring points and the peak is then: $x_{i-1} < x_i > x_{i+1}$. The gradient-based method evaluates the first derivative at the point that is equal to zero if the point is a local maximum, local minimum, or inflection point~\cite{adikaram_non-parametric_2016}.

@@

\subsection{Time-frequency domain features}
The \textbf{Short-time Fourier transform} (STFT) splits the time-domain signal into equal-length intervals. Individual chunks have a 50\% overlap and are multiplied with weights of window function to balance scalloping loss and spectral leakage due to the Fourier transform periodicity assumption. Traditionally, the Hann window is commonly used instead of the rectangular window~\cite{ziaran_technicka_2013,noauthor_iso_2016_2}.

@@

In washing machine status classification, the discrete wavelet transform with Daubechies wavelet (db4) and fifth-level decomposition provides features with a combination of approximation (cA5) and detail coefficients (cD1, \dots, cD5). Washing machines belong to three categories: no fault, electric motor clamping screws problem, and a loose or broken counterweight. Extracted measures were sample mean and sample variances over autocorrelation functions of coefficients (AcD$n$) and smoothed coefficients cD1, cD2 by moving average filter~\cite{goumas_classification_2002}.

@@

Tool wear diagnosis based on acoustic emission (AE) signal considers \emph{wavelet packet energy} in bands $E_{8}$, $E_{10}$, $E_{12}$, \emph{energy ratios} $P_{8}$, $P_{13}$, and \emph{energy entropy} as having high correlation ($|r| > 0.8$) with the band saw flank face width. The acoustic signal is decomposed into three layers using Daubechies db3 wavelet. The bottom layer contains bands numbered 7 through 14, each with a bandwidth of 62.5 kHz, because of the 1 MHz sampling frequency. The feature vector constructed in the article includes other statistical metrics out of power spectral density that have reached a notable correlation with evolving wear. The statistics are \emph{skewness}, \emph{kurtosis}, \emph{shape factor}, and \emph{centroid frequency}~\cite{zhuo_research_2022}.

@@

Wavelet domain features and associated spectral segmentation appear to be a powerful tool for conserving data bandwidth. These variables turn out to be difficult to communicate and comprehend. Therefore, we further use more standardized metrics.

@@

\item \textbf{Correlation coefficient} - expresses the linear relationship between two variables. The codependent variables are of three sorts: quantitative, ordinal, and nominal. The choice of coefficient calculation is determined by the type of variables under consideration as shown in Table \ref{tab:corr-coef}.

@@

\item \textbf{Mutal information} - quantifies the dependence between features, or between features and class labels. It is almost identical to Information Gain. The probability distribution of proximity of variables derives from the relative entropy known as the Kullback-Leibler distance. Mutual information presumes variables are discrete. In the case of quantitative variables, mutual information is estimated by binning or nearest neighbours methods~\cite{ross_mutual_2014}.

@@

The measurements coming in the steaming fashion have to be processed in a single pass. The detection model must deal with the minimal admissible assumptions about the nature of the input events. The outliers are derived based on non-parametric statistical models, nearest-neighbour clustering, and isolation-based approaches~\cite{gervasi_anomaly_2020}.
\bigbreak

\textbf{DenStream} is a density-based algorithm adapted from DBSCAN to cluster streaming data of arbitrarily shaped groups. Samples it includes in the first step into coherent clusters are core data points in each other's neighbourhoods. Core points have at least \emph{MinPts} ($\mu$) points in their neighbourhood of radius \emph{Eps} ($\varepsilon$) units. Then non-core points in the proximity area of the core point are attached to the cluster containing it~\cite{aggarwal_data_2014}.

Quality of clustering results is evaluated by \emph{Silhouette score} in the range [-1; 1]. It demands points within clusters to have high cohesion and at the same time to have large separation from other clusters. The low score indicates a too-small or too-large number of clusters~\cite{rousseeuw_rousseeuw_1987}.

@@

In the online maintenance phase, DenStream summarizes the nearby observations into core \emph{micro-clusters} that can be potential micro-clusters or outlier \emph{micro-clusters} (Fig.~\ref{fig:denstream})~\cite{ghesmoune_state---art_2016}. The (outlier) \emph{o-micro-clusters} can grow into (potential) \emph{p-micro-clusters} when they encompass $\beta \mu$ points. The outliers are discounted after some time in accordance with the decay function: $f(t) = 2^{-\lambda t}$ or below the lower weight limit $\xi$. The on-demand offline stage runs DBSCAN over the approximate representation in micro-clusters to deliver final apportionment~\cite{cao_density-based_2006}.

\subsection{Classification}
Accurate multi-class classification of machine faults according to the characteristics of known causes is a much more difficult task than novelty detection. Fault combinations have to be recorded and transformed into feature space. Interactions among fault root causes have to be considered. We are aware of rapid advances in knowledge transfer for deep neural networks~\cite{maurya_condition-based_2021}. So far, solutions seem to be hard to implement. Therefore, we opt to use a simpler model.

The performance of classification is estimated by several metrics on the validation set obtained using hold-out or cross-validation techniques. Frequently used quantities for classifier model evaluation include accuracy, precision, recall, f1 score, area under the ROC curve, and counts of hits and misses in a confusion matrix.
\bigbreak

\textbf{K-nearest neighbours} (k-NN) assigns the data point to the class where the majority of $k$ closest instances belong (Fig.~\ref{fig:KNN}). It means it can work in a semi-supervised environment because it can infer labels just from knowing a few annotations. The major drawback of k-NN is a preference for the majority class in imbalanced class-size datasets. The issue is mitigated with class weights or resampling classes by oversampling or undersampling. The model without class balancing drops in precision from 15\% to over 40\% depending on the imbalance ratio~\cite{shi_improving_2020}. The k-NN algorithm requires features to be normalized to assign the same importance to each predictor.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/analysis/KNN.png}
        \caption{k-NN with k = 5}
        \label{fig:KNN}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/analysis/M-tree.png}
        \caption{M-tree data structure}
        \label{fig:m-tree}
    \end{subfigure}
    \caption{Nearest neighbours classification algorithm~\cite{chen_skyline_2009}}
\end{figure}

The meaning of distance between feature vectors $\mathbf{x}$, $\mathbf{y}$ have to be defined, so several metrics are available like \emph{Euclidian distance}, \emph{Mahalanobis distance}, \emph{Manhatann distance} etc. (Tab.~\ref{tab:KNN-distance})~\cite{sheng_review_2020, abu_alfeilat_effects_2019}. The optimal $k$ parameter is set by supervised learning according to the breaking point in the elbow curve that plots choices of $k$ against the error rate. The demanding neighbourhood queries are sped up using spatial indexes in databases that utilize search trees such as kd-tree, R-tree, or M-tree.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|}
\hline
\textbf{Distance}     & \textbf{$d(\mathbf{x}, \mathbf{y})$}                                   \\ \hline
Manhattan distance	 & $ |x_i - y_i| $													   \\ \hline
Euclidian distance    & $ \sqrt{\sum_{i = 1}^{n}(x_i - y_i)^2} $                               \\ \hline
Mahalanobis distance  & $ (\mathbf{x} - \mathbf{y})^T C^{-1} (\mathbf{x} - \mathbf{y}) $       \\ \hline
\end{tabular}
\caption{Distance metrics for k-NN}
\label{tab:KNN-distance}
\end{table}
